---
title: "Final paper code consolidated"
author: "Jialu Li"
date: "6/11/2018"
output: html_document
---
```{r}
library(lubridate)
library(magrittr)
library(dplyr) 
library(ggplot2)
library(sp)
library(maptools)
library(rgdal)
library(raster)
library(plotly)
library(viridis)
library(knitr)
library(DT)
library(rgdal)
library(scales)
library(RColorBrewer)
library(ggthemes)
library(data.table)
library(anytime)
```


#Data Visualization
##load data files
```{r}
getwd()
uber_apr_14 <- read.csv("uber-raw-data-apr14.csv")
uber_may_14 <- read.csv("uber-raw-data-may14.csv")
uber_jun_14 <- read.csv("uber-raw-data-jun14.csv")
uber_jul_14 <- read.csv("uber-raw-data-jul14.csv")
uber_aug_14 <- read.csv("uber-raw-data-aug14.csv")


#UberJanJune15 <- read.csv("uber-raw-data-janjune-15.csv")

head(uber_apr_14)
```

Binding all the Data into one
```{r}
uber_2014 <- rbind(uber_apr_14, uber_may_14,uber_jun_14,uber_jul_14,uber_aug_14)
#uber_2014$Date.Time <- anytime(as.character(uber_2014$Date.Time))

uber_2014$Date.Time <- as.character(uber_2014$Date.Time)

uber_2014$Date.Time = as.POSIXct(uber_2014$Date.Time,format="%m/%d/%Y %H:%M")

data.table(uber_2014,nrow=30)

class(uber_2014$Date.Time)

uber_2014$Day <- factor(day(uber_2014$Date.Time))
uber_2014$Month <- factor(month(uber_2014$Date.Time))
uber_2014$Year <- factor("2014")
uber_2014$Weekday <- factor(wday(uber_2014$Date.Time))
uber_2014$Hour <- factor(hour(uber_2014$Date.Time))


data.table(head(uber_2014))
data.table(uber_2014,nrow=30)
```

##No. of Uber Trips by Day made in New York City
```{r}
cols <- brewer.pal(8, "Paired")
pal <- colorRampPalette(cols)
trips_day <- group_by(uber_2014, Day) %>% summarise(Trips = n())
data.table(trips_day)
ggplot(trips_day, aes(Day, Trips)) + geom_bar(stat = "identity", fill = "maroon") +
    ggtitle("No. of Uber Trips by Day made in New York City")
```

##No. of Uber Trips by Day & Month made in New York City
```{r}
trips_day_month <- group_by(uber_2014, Day, Month) %>% summarise(Trips = n())
ggplot(trips_day_month, aes(Day, Trips, fill = Month)) + geom_bar(stat = "identity") +
    ggtitle("No. of Uber Trips by Day & Month made in New York City") + scale_y_continuous() +
      scale_fill_manual(values = pal(6)) + facet_grid(Month ~ .)

```

No. of Uber Trips Happening in New York City Every Hour
```{r}
library(DT)
library(ggplot2)
cols <- brewer.pal(8, "Dark2")
pal <- colorRampPalette(cols)
trips_hour <- group_by(uber_2014, Hour) %>% summarise(Trips = n())
data.table(trips_hour)
ggplot(trips_hour, aes(Hour, Trips)) + geom_bar(stat = "identity", fill = "orange") + ggtitle("No. of Uber Trips Happening in New York City Every Hour")
trips_hour_month <- group_by(uber_2014, Hour, Month) %>% summarise(Trips = n())
ggplot(trips_hour_month, aes(Hour, Trips, fill = Month)) + geom_bar(stat = "identity") + ggtitle("No. of Uber Trips Happening in New York City Every Hour for Different Months") + scale_y_continuous() + scale_fill_manual(values = pal(6)) + facet_grid(Month ~ .)
```


##Uber Trips made According to The Days in New York City
```{r}
cols <- brewer.pal(8, "Spectral")
pal <- colorRampPalette(cols)
trips_wday <- group_by(uber_2014, Weekday) %>% summarise(Trips = n())

datatable(trips_wday)
ggplot(trips_wday, aes(Weekday, Trips)) + geom_bar(stat = "identity", fill = "Darkred") + ggtitle("No. of Uber Trips Happening in New York City For Different Days of the Week")
trips_wday_month <- group_by(uber_2014, Weekday, Month) %>% summarise(Trips = n())
ggplot(trips_wday_month, aes(Weekday, Trips, fill = Month)) + geom_bar(stat = "identity", position = "dodge") + ggtitle("No. of Uber Trips Happening in New York City for WeekDays of Different Months") + scale_y_continuous() + scale_fill_manual(values = pal(6))
```


Heatmap for Uber Trips by BASE and Day
```{r, echo = FALSE}
trips_base_wday <- group_by(uber_2014, Base, Weekday) %>% summarise(Trips = n())
ggplot(trips_base_wday, aes(Base, Weekday, fill = Trips)) + ggtitle("Heatmap of Uber Trips by BASE and Day") + geom_tile(color = "white")
```


#Pickup number comparison
```{r}
getwd()
Uber_2015_raw <- read.csv("uber-raw-data-janjune-15.csv")

Uber_2015_raw$Pickup_date = as.character(Uber_2015_raw$Pickup_date)
Uber_2015_raw$Pickup_date_new = as.POSIXct(Uber_2015_raw$Pickup_date,format="%Y-%m-%d %H:%M")

class(Uber_2015_raw$Pickup_date_new)


Uber_2015_raw$Month <- month(Uber_2015_raw$Pickup_date_new)

head(Uber_2015_raw)
nrow(Uber_2015_raw)

#data.table(head(Uber_2015_raw))

```

UBER: only select 2015 Jan and June month data to speed up
```{r}

Uber_2015_test_short= subset(Uber_2015_raw,Uber_2015_raw$Month == "1"|Uber_2015_raw$Month =="6" )
nrow(Uber_2015_test_short)

Uber_2015_test_short$Year = as.numeric("2015")

head(Uber_2015_test_short) 

```

Get Uber 2015 April subset
```{r}
Uber_2015_Apr= subset(Uber_2015_raw,Uber_2015_raw$Month == "4" )

Uber_2015_Apr$Hour <- as.numeric(factor(hour(Uber_2015_Apr$Pickup_date_new)))

head(Uber_2015_Apr)

#define rush hour as between 17pm-20pm
Uber_2015_Apr_rushhour = subset(Uber_2015_Apr,Uber_2015_Apr$Hour>=17&Uber_2015_Apr$Hour<=20)
Uber_2015_Apr_rushhour $Year = as.numeric("2014")
head(Uber_2015_Apr_rushhour)
```

Tally the rush hour
```{r}
Uber_2015_Apr_by_month_hour <- Uber_2015_Apr_rushhour %>% 
    group_by(Year,Month, Hour) %>% 
    tally() %>% 
    arrange(Year,Month, Hour)
data.table(Uber_2015_Apr_by_month_hour)

setnames(Uber_2015_Apr_by_month_hour,"n","2015_Uber_Apr_pickup_number")
```

Import 2014 April dataset
```{r}
head(uber_apr_14)
class(uber_apr_14$Date.Time)

uber_apr_14$Date.Time <- as.character(uber_apr_14$Date.Time)

uber_apr_14$Date.Time.New <- as.POSIXct(uber_apr_14$Date.Time,format="%m/%d/%Y %H:%M")

uber_apr_14$Month <- month(uber_apr_14$Date.Time.New)
uber_apr_14$Hour <- as.numeric(factor(hour(uber_apr_14$Date.Time.New)))

uber_apr_14$Year = as.numeric("2014")
```
define rush hour as between 17pm-20pm
```{r}
uber_apr_14_rushhour = subset(uber_apr_14,uber_apr_14$Hour>=17&uber_apr_14$Hour<=20)
head(uber_apr_14_rushhour)
```

Tally the 2014 April rush hour
```{r}
uber_apr_14_by_month_hour <- uber_apr_14_rushhour %>% 
    group_by(Year,Month, Hour) %>% 
    tally() %>% 
    arrange(Year,Month, Hour)
data.table(uber_apr_14_by_month_hour)

setnames(uber_apr_14_by_month_hour,"n","2015_Uber_Jun_pickup_number")
```

merge the 2 dataset
```{r}
rush_hour_comparison= merge(uber_apr_14_by_month_hour,Uber_2015_Apr_by_month_hour,by="Hour",all=FALSE )
data.table(rush_hour_comparison)
```


Import 2014 June dataset
```{r}
head(uber_jun_14)
class(uber_jun_14$Date.Time)

uber_jun_14$Date.Time <- mdy_hms(uber_jun_14$Date.Time)

uber_jun_14$Date.Time <- as.Date(uber_jun_14$Date.Time,format = "%m/%d/%Y")

uber_jun_14$Month <- month(ymd(uber_jun_14$Date.Time))

uber_jun_14$Year = as.numeric("2014")
```

UBER: Count 2015 Uber pickup Number of Trips by Month
```{r} 
library(dplyr)

by_month_year <- Uber_2015_test_short %>% 
    group_by(Month, Year) %>% 
    tally() %>% 
    arrange(Year, Month)
data.table(by_month_year)
```


```{r}
by_month_year_2014jun <- uber_jun_14 %>% 
    group_by(Month, Year) %>% 
    tally() %>% 
    arrange(Year, Month)

data.table(by_month_year_2014jun)
```

```{r}
by_month_year_all= rbind(by_month_year_2014jun, by_month_year)

by_month_year_all= subset(by_month_year_all[2:4,])
by_month_year_all
```

UBER: Count 2015 Uber pickup Number of Trips by Date
```{r} 
library(dplyr)

by_date_year <- Uber_2015_test_short %>% 
    group_by(Pickup_date, Year) %>% 
    tally() %>% 
    arrange(Year, Pickup_date)
head(data.table(by_date_year))
```


plot the histogram of Uber pickup number
```{r}
qplot(by_date_year$n, geom="histogram") 

qplot(by_date_year$n,
      geom="histogram",
      binwidth = 1,  
      main = "Histogram for 2015 Uber pickup number distribution", 
      xlab = "Pickup number",  
      fill=I("blue"), 
      col=I("red"), 
      alpha=I(.2),
      xlim=c(0,125000))
```





```{r}

uber_2015_pickupnumber <- read.csv("2015 Uber pickup data by date.csv")
weather_2015 <- read.csv("2015 weather data.csv")


uber_2015_pickupnumber$Date_2015 = as.Date(uber_2015_pickupnumber$Date_2015, format = "%m/%d/%Y")

head(uber_2015_pickupnumber,nrow=40)

weather_2015$Date_2015=as.Date(weather_2015$Date_2015, format = "%m/%d/%y") 
head(weather_2015)

```



merge data 
```{r}
New_2015=merge(uber_2015_pickupnumber,weather_2015, by="Date_2015",all=FALSE)

head(New_2015)  

summary(New_2015)
str(New_2015)
```

#convert precipitation into categorical event
```{r}

class(New_2015$Precipitation_inch)

New_2015$Precipitation_inch= as.numeric(as.character(New_2015$Precipitation_inch))

head(New_2015)

```

plot to analyze whether rainy days have more pickup
conclusion: during rainy days have more pickup 
```{r}
ggplot(New_2015, aes(x=Rain_Event, y=Number_of_pickup)) +geom_boxplot()+ ggtitle("Number_of_pickup vs. Rain_Event") 

```

## two-way ANOVA
p value = 0.014, i can reject null hypothesis that rainy days don't have higher pickup number
```{r}
results= lm(Number_of_pickup ~ Rain_Event, data=New_2015)
anova(results)

```

check relevant model assumptions. This includes making QQ-plots and residual plots. 
variance:doesn't violate the assumption that they need to have equal variances
our assumptions about variance and normality are not violated 
```{r}
par(mfrow=c(2,2))
plot(results)
```



linear modeling
```{r}
lm_temp=lm(Number_of_pickup ~ Temp_avg, data = New_2015)
summary(lm_temp)
plot(Number_of_pickup ~ Temp_avg, data = New_2015)+abline(lm_temp,lwd=2)

```

Lm with precipitation
```{r}
lm_precip= lm(Number_of_pickup~ Precipitation_inch,data = New_2015)
summary(lm_precip)
```


```{r}
lm_temp_3=lm(Number_of_pickup ~ Temp_avg+ Precipitation_inch+Temp_avg:Precipitation_inch, data = New_2015)
summary(lm_temp_3)
```



##Analysis of covariance: ANCOVA: One continuous variable and one discrete variable
```{r}
ANCOVA_model = lm(Number_of_pickup ~ Temp_avg+Rain_Event, data = New_2015)
anova(ANCOVA_model)
summary(ANCOVA_model)



library(car)
anova(multi_model, type = "III")
```

```{r}
par(mfrow=c(2,2))
plot(ANCOVA_model)
```


try interaction between 2 variables:
```{r}
ANCOVA_model_2 = lm(Number_of_pickup ~ Temp_avg*Rain_Event, data = New_2015)
anova(ANCOVA_model_2)
```

comparing the 2 models, there is no significant difference between the 2 models.
```{r}
anova(ANCOVA_model, ANCOVA_model_2)
```

#Fit polynomial regression
```{r}
poly_2= lm(Number_of_pickup ~ poly(Temp_avg, 2),data = New_2015)  
summary(poly_2)

poly_3= lm(Number_of_pickup ~ poly(Temp_avg, 3),data = New_2015)  
summary(poly_3)
```

plot 
```{r}
library(ggplot2)
ggplot(New_2015,aes(Temp_avg, Number_of_pickup))+ geom_point() + labs(title = ('Number_of_pickup vs Temp_avg Polynomial Regressions'),x = 'Temp_avg',yl = "Number_of_pickup")+ geom_line(data = fortify(poly_2),aes(x=New_2015$Temp_avg, y=.fitted),col='blue')+ geom_line(data = fortify(poly_3),aes(x=New_2015$Temp_avg, y=.fitted),col='green') + geom_smooth(method = 'auto') 

```

#Explore correlation relationship with average speed
Merge weather with speed data 
```{r}

YC_Jan_June_merged=rbind(YC_2015_Jan_short,YC_2015_June)
head(YC_Jan_June_merged)

YC_Jan_June_merged$Date_2015 = as.Date(as.character(YC_Jan_June_merged$New_pickup_datetime, format = "%m/%d/%Y","EST"),format = "%m/%d/%Y")

head(YC_Jan_June_merged)
head(weather_2015)

YC_Jan_June_merged_new=merge(YC_Jan_June_merged,weather_2015, by="Date_2015",all=FALSE)

head(YC_Jan_June_merged_new,n=30)

YC_Jan_June_merged_new$Temp_avg=as.numeric(YC_Jan_June_merged_new$Temp_avg)
YC_Jan_June_merged_new$Temp_low=as.numeric(YC_Jan_June_merged_new$Temp_low)
YC_Jan_June_merged_new$Precipitation_inch=as.numeric(YC_Jan_June_merged_new$Precipitation_inch)
YC_Jan_June_merged_new$Weekday <- factor(wday(YC_Jan_June_merged_new$Date_2015))

head(YC_Jan_June_merged_new)

```

??Get average speed by week day: but unable to retrieve it??
```{r}
speed_month <- group_by(YC_Jan_June_merged_new,Weekday) %>% mean(speed)
ggplot(trips_hour_month, aes(Hour, Trips, fill = Month)) + geom_bar(stat = "identity") + ggtitle("No. of Uber Trips Happening in New York City Every Hour for Different Months") + scale_y_continuous() + scale_fill_manual(values = pal(6)) + facet_grid(Month ~ .)
```

```{r}
plot(speed ~ Rain_Event, data=YC_Jan_June_merged_new)

ggplot(YC_Jan_June_merged_new, aes(x=Rain_Event, y=speed)) + geom_boxplot()+ coord_cartesian(ylim = c(0, 90))

```

```{r}
ANOVA_results_2= aov(speed ~ Rain_Event, data=YC_Jan_June_merged_new)
summary(ANOVA_results_2)
```

```{r}
par(mfrow=c(2,2))
plot(ANOVA_results_2)
```

```{r}
t.test(speed ~Rain_Event, data=YC_Jan_June_merged_new)
```


#Uni-variate linear modeling
```{r}
YC_lm_temp=lm(speed ~ Temp_avg, data = YC_Jan_June_merged_new)
summary(YC_lm_temp)

plot(speed ~ Temp_avg, data = YC_Jan_June_merged_new)+abline(YC_lm_temp,lwd=2)
```

```{r}
YC_lm_precip=lm(speed ~ Precipitation_inch, data = YC_Jan_June_merged_new)
summary(YC_lm_precip)


plot(speed ~ Precipitation_inch, data = YC_Jan_June_merged_new)+abline(YC_lm_precip,lwd=2)
```

##check correlation of interaction term
```{r}
YC_lm_interaction = lm(speed ~ Temp_avg*Precipitation_inch, data = YC_Jan_June_merged_new)
summary(YC_lm_interaction)
```

```{r}
myvars = c("Temp_avg", "Precipitation_inch")

library(RColorBrewer)
library(ggplot2)
library(lattice)

pairs(YC_Jan_June_merged_new[myvars], pch = 15)
```


##check Pearson correlation
```{r}
correlation_1=cor.test(YC_Jan_June_merged_new$speed,YC_Jan_June_merged_new$Temp_avg,method='pearson')
correlation_1
```


```{r}
correlation_2=cor.test(YC_Jan_June_merged_new$speed,YC_Jan_June_merged_new$Precipitation_inch,method='pearson')
correlation_2

#the result is counter-intuitive because Precipitation_inch and Temp_avg do not have normal distribution. 
```

##MAKE pairwise scatterplot
```{r}
library(GGally)
ggpairs(YC_Jan_June_merged_new[,c("speed", "Temp_avg", "Precipitation_inch")])
```

